\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-28 notes/2loopNLSM.tex
small edit
CS edits

@@

We begin with an overview of color-kinematics duality in
\cref{background}, and define the notion of ``globally'' color-dual integrands in \cref{sec:bootstrap}. 
We then construct a manifestly color-dual theory in
\cref{sec:cubic}, which generates $D$-dimensional color-dual $n$-point
numerators for both NLSM and the self-dual sector of Yang-Mills
through one-loop. The construction of this theory, which we dub
\textit{semi-abelian Yang-Mills}, relies on isolating the cubic sector
of pure Yang-Mills amplitudes from the four-point contacts that are
needed to fully realize non-abelian gauge symmetry. Despite the
potency of this theory at one-loop, the construction runs into an
obstruction at two-loop, which we describe in \cref{2LoopObstruction}.

@@

Thus for our approach at one-loop, we aim to build color-dual integrands directly from $D$-dimensional kinematic factors that are $\mathcal{O}((\varepsilon\varepsilon)^1)$ at tree-level, and $\mathcal{O}((\varepsilon\varepsilon)^0)$ at one-loop. The $D$-dimensional organizational principle underlying this construction can be understood in terms of a vector amplitude decomposition introduced by one of the authors \cite{Pavao:2022kog},
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
where we have introduced the shorthand notation, $\varepsilon_{(ij)\cdots (kl)} \equiv (\varepsilon_i \varepsilon_j)\cdots (\varepsilon_k \varepsilon_l)$, and  $S^{2|k}_{\sigma}$ is the set of $k$ pairs of external legs appearing in the color-ordered label list, $\sigma$. For example, at four-point, $S^{2|1}_{\sigma} =  \{(12),(13),(14),(23),(24),(34)\}$, and $S^{2|2}_{\sigma} =  \{(12)(34),(13)(24),(14)(23)\}$. The sum thus selects out kinematic building blocks, $\Delta_{(\sigma)}^{(\rho)}$,that are each weighted by different polarization dot products. Some simple four-point examples include, 
\be
\Delta^{(13)}_{(1234)} =\frac{(k_1 \varepsilon_2)(k_{3} \varepsilon_4)}{s_{12}}+\frac{(k_1 \varepsilon_4)(k_{3} \varepsilon_2)}{s_{14}}, \qquad \Delta^{(13)(24)}_{(1234)} =1, \qquad \Delta^{(12)(34)}_{(1234)} =\frac{s_{13}}{s_{12}}\,. 
\ee
We direct the reader to ref.~\cite{Pavao:2022kog} for further details. The expansion of \cref{eq:pureVecRABD} comes with the added advantage of making the transmutation relations of \cite{Cheung:2017ems} absolutely manifest. By a simple mass-dimension argument \cite{ElvangHuangReview}, we know the tree-level cubic sector of Yang-Mills must be at
$\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic-YM}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to
$\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot
products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic-YM}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks
must obey a set of Ward-identities between different ``helicity'', or
$(\varepsilon \varepsilon)^n$, sectors in order for the full amplitude
to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of
the cubic sector for Yang-Mills, and will demonstrate that the
resulting amplitudes with manifestly color-dual Feynman rules are
equivalent to both SDYM and NLSM through one-loop.

@@

Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.

@@

In addition to the tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands. After plugging in suitable on-shell states, the integrands
produced by semi-abelian YM correspond to those of SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}.  Due to the
kinetic mixing that stems from the broken Bose symmetry, \cref{2loopTension} shows why it is
impossible to construct a two-loop integrand for semi-abelian YM theory. In \cref{2loopBoot}, we overcame this obstruction by successfully constructing two-loop NLSM numerators with an ansatz based bootstrap. However, we found that the same approach fails for two-loop pure Yang-Mills, and pin-pointed the failure to a particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry, and in this paper we have argued that off-shell locality\footnote{An example of this can already be found in the literature, where locality is relaxed in Ref.~\cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.} must
be abandoned for color-dual constructions of multi-loop Yang-Mills. To expand the available function space, we propose that one should consider including rational functions of kinematics, rather than just polynomials.

@@

Below we review the four manifestly color-dual theories that we mention in
the text, self-dual Yang-Mills (SDYM) \cite{Monteiro2011pc}, two formulations of NLSM \cite{Cheung:2016prv,Cheung:2020djz,Cheung:2021zvb}, and ZM \cite{Zakharov:1973pp, Cheung:2022mix}.  All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM}, which gives the cubic sector of pure YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint directly on the
field strength. In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Since Bose symmetry is broken, the ensuing kinetic mixing prevents color-dual representations beyond one-loop without the introduction of additional interactions.

@@

\subsection{J-theory}
$J$-theory is obtained through a first order reformulation of NLSM in terms of the chiral current \cite{Cheung:2021zvb}.\footnote{For related formulations of NLSM that do not
treat color-kinematics duality see \cite{Freedman:1980us,
  Slavnov:1971mz}.} Two constraints are placed on the chiral current.  First, the field strength associated with $J_\mu$ should vanish,
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0\, ,
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge,
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0\, ,
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
Since both theories are in Lorenz gauge this means that \emph{J-theory and semi-abelian YM are identical}. Furthermore, this theory is one-to-one with the observables amplitudes of $YZ$-theory when only considering on-shell $Z$$\bar{Z}$ states. As such, the single vertex in this theory is the same as the pure vector vertex of $YZ$ theory in Lorenz gauge,
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)\, .
\end{equation}
By calculating an off-shell four-point correlation function it is possible to check that the Lagrangians of $J$-theory and semi-YM theory automatically satisfies color-kinematics duality, as described in \cref{offShellCK}.
As might be expected, the kinematic algebra is the diffeomorphism algebra \cite{Cheung:2021zvb}. Note that the different external particles break manifest Bose symmetry.

@@

@@

We use the same conventions as Ref.~\cite{jjmcTASI2014}, which
we quote now. The component-wise definitions of the spinor brackets are
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\end{align}
where the $a_i$ are components of the four-vector,
$k^\mu_a = (a_0,a_1,a_2,a_3)$. For massless momenta, $k_a$ and $k_b$,
one can apply the on-shell condition, $k_a^2=k_b^2=0$, to show that
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,.
\end{equation}
Four-dimensional polarization dot products with fixed helicity states
are mapped to
\begin{equation}\label{eq:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}\,,
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}\,,
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} \,,
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 \,,
\end{aligned}
\end{equation}
Note that given the above definition, the spinor helicity variables
carry the same mass dimension for both the angle and square
brackets. In terms of the light-cone coordinates that we employed in
\cref{sec:CKLagrangians}, the angles and squares can be rewritten as
\begin{align}
\langle ab \rangle &= \frac{a_w b_u-a_u b_w}{\sqrt{a_ub_u}}\,,
\\
[ab] &= \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{\sqrt{a_ub_u}}\,.
\end{align}
The conventions we use in the text where the holomorphic spinor,
$X(p,k)$, carries extra mass dimension amounts to shifting
the bracket by an overall factor,
\begin{align}
\langle ab \rangle &\rightarrow a_w b_u-a_u b_w \equiv X(a,b)\,,
\\
[ab] &\rightarrow \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{a_ub_u}\equiv Q(a,b)\, .
\end{align}
With this choice we still obtain the same completeness relation for the SDYM spinors,
\begin{equation}
X(a,b)Q(b,a) = \frac{1}{2}s_{ab}\, .
\end{equation}

>>>>

\newcommand{\jm}[1]{\textcolor{blue}{\textbf{JM: }{#1}}}

@@

We begin with an overview of color-kinematics duality in
\cref{background}, and define the notion of ``globally'' color-dual integrands in \cref{sec:bootstrap}. 
We then construct a manifestly color-dual theory in
\cref{sec:cubic}, which generates $D$-dimensional color-dual $n$-point
numerators for both NLSM and the self-dual sector of Yang-Mills
through one-loop. \jm{Mention CS here?} The construction of this theory, which we dub
\textit{semi-abelian Yang-Mills}, relies on isolating the cubic sector
of pure Yang-Mills amplitudes from the four-point contacts that are
needed to fully realize non-abelian gauge symmetry. Despite the
potency of this theory at one-loop, the construction runs into an
obstruction at two-loop, which we describe in \cref{2LoopObstruction}.

@@

Thus for our approach at one-loop, we aim to build color-dual integrands directly from $D$-dimensional kinematic factors that are $\mathcal{O}((\varepsilon\varepsilon)^1)$ at tree-level, and $\mathcal{O}((\varepsilon\varepsilon)^0)$ at one-loop. The $D$-dimensional organizational principle underlying this construction can be understood in terms of a vector amplitude decomposition introduced by one of the authors \cite{Pavao:2022kog},
\begin{equation}\label{eq:pureVecRABD}
A_{(\sigma)}^{\text{YM}} = \sum_{k=0}^{\lfloor |\sigma|/2\rfloor}\sum_{\rho \in S^{2|k}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
where we have introduced the shorthand notation, $\varepsilon_{(ij)\cdots (kl)} \equiv (\varepsilon_i \varepsilon_j)\cdots (\varepsilon_k \varepsilon_l)$, and  $S^{2|k}_{\sigma}$ is the set of $k$ pairs of external legs appearing in the color-ordered label list, $\sigma$. For example, at four-point, $S^{2|1}_{\sigma} =  \{(12),(13),(14),(23),(24),(34)\}$, and $S^{2|2}_{\sigma} =  \{(12)(34),(13)(24),(14)(23)\}$. The sum thus selects out kinematic building blocks, $\Delta_{(\sigma)}^{(\rho)}$,that are each weighted by different polarization dot products. Some simple four-point examples include, 
\be
\Delta^{(13)}_{(1234)} =\frac{(k_1 \varepsilon_2)(k_{3} \varepsilon_4)}{s_{12}}+\frac{(k_1 \varepsilon_4)(k_{3} \varepsilon_2)}{s_{14}}, \qquad \Delta^{(13)(24)}_{(1234)} =1, \qquad \Delta^{(12)(34)}_{(1234)} =\frac{s_{13}}{s_{12}}\,. 
\ee
We direct the reader to Ref.~\cite{Pavao:2022kog} for further details. The expansion of \cref{eq:pureVecRABD} comes with the added advantage of making the transmutation relations of \cite{Cheung:2017ems} absolutely manifest. By a simple mass-dimension argument \cite{ElvangHuangReview}, we know the tree-level cubic sector of Yang-Mills must be at
$\mathcal{O}((\varepsilon\varepsilon)^1)$,
\begin{equation}
A_{(\text{tree})}^{\text{cubic-YM}} = \sum_{\rho \in S^{2|1}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\, ,
\end{equation}
while at one-loop, the cubic sector corresponds to
$\mathcal{O}((\varepsilon\varepsilon)^0)$ in polarization dot
products,
\begin{equation}
A_{(\text{1-loop})}^{\text{cubic-YM}} = \sum_{\rho \in S^{2|0}_{\sigma}}\varepsilon_{(\rho)} \Delta_{(\sigma)}^{(\rho)}\,.
\end{equation}
When decomposed in this way, the polarization stripped building blocks
must obey a set of Ward-identities between different ``helicity'', or
$(\varepsilon \varepsilon)^n$, sectors in order for the full amplitude
to be gauge invariant,
 \begin{equation}\label{eq:GIrelA}
\Delta_{(\sigma)}^{(\rho)}\Big|_{\epsilon_i\rightarrow k_i} =-
\sum_{j \in \rho^c} (k_i \varepsilon_j)\Delta_{(\sigma )}^{(\rho\cup (ij))} \,.
\end{equation}
With this in hand, we can construct a Lagrangian description of the of
the cubic sector for Yang-Mills, and will demonstrate that the
resulting amplitudes with manifestly color-dual Feynman rules are
equivalent to both SDYM and NLSM through one-loop.

@@

Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, self-dual Yang-Mills
\cite{Monteiro2011pc}, and Chern-Simons theory \cite{Ben-Shahar:2021zww}. \jm{I mentioned CS.}  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}\label{semiYMvert}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be\label{semiYMprop}
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.

@@

In addition to the tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands. After plugging in suitable on-shell states, the integrands
produced by semi-abelian YM correspond to those of SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}.  \jm{Mention CS?  Maybe something like `SDYM in (), NLSM in (), and Chern-Simons theory in \cref{sec:CSTheory}.'?}  Due to the
kinetic mixing that stems from the broken Bose symmetry, \cref{2loopTension} shows why it is
impossible to construct a two-loop integrand for semi-abelian YM theory. In \cref{2loopBoot}, we overcame this obstruction by successfully constructing two-loop NLSM numerators with an ansatz based bootstrap. However, we found that the same approach fails for two-loop pure Yang-Mills, and pin-pointed the failure to a particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry, and in this paper we have argued that off-shell locality\footnote{An example of this can already be found in the literature, where locality is relaxed in Ref.~\cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.} must
be abandoned for color-dual constructions of multi-loop Yang-Mills. To expand the available function space, we propose that one should consider including rational functions of kinematics, rather than just polynomials.

@@

Below we review the five manifestly color-dual theories that we mention in
the text, self-dual Yang-Mills (SDYM) \cite{Monteiro2011pc}, two formulations of NLSM \cite{Cheung:2016prv,Cheung:2020djz,Cheung:2021zvb}, ZM \cite{Zakharov:1973pp, Cheung:2022mix}, and Chern-Simons theory \cite{Ben-Shahar:2021zww}.  \jm{I added CS} All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM}, which gives the cubic sector of pure YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint directly on the
field strength. In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Since Bose symmetry is broken, the ensuing kinetic mixing prevents color-dual representations beyond one-loop without the introduction of additional interactions.

@@

\subsection{$J$-theory}
$J$-theory is obtained through a first order reformulation of NLSM in terms of the chiral current \cite{Cheung:2021zvb}.\footnote{For related formulations of NLSM that do not
treat color-kinematics duality see \cite{Freedman:1980us,
  Slavnov:1971mz}.} Two constraints are placed on the chiral current.  First, the field strength associated with $J_\mu$ should vanish,
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0\, ,
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge,
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0\, ,
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
Since both theories are in Lorenz gauge this means that \emph{$J$-theory and semi-abelian YM are identical}. Furthermore, this theory is one-to-one with the observables amplitudes of $YZ$-theory when only considering on-shell $Z$$\bar{Z}$ states. As such, the single vertex in this theory is the same as the pure vector vertex of $YZ$ theory in Lorenz gauge,
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{2})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{1})(\varepsilon_1\bar{\varepsilon}_3)\, .
\end{equation}
By calculating an off-shell four-point correlation function it is possible to check that the Lagrangians of $J$-theory and semi-YM theory automatically satisfies color-kinematics duality, as described in \cref{offShellCK}.
As might be expected, the kinematic algebra is the diffeomorphism algebra \cite{Cheung:2021zvb}. Note that the different external particles break manifest Bose symmetry.

@@

\subsection{Chern-Simons theory}
\label{sec:CSTheory}
The final theory that is unified by the semi-abelian Yang-Mills construction is \jm{(tree level) ?} Chern-Simons (CS) theory. As we will show below, plugging in appropriate on-shell states to the semi-YM numerators that we constructed in the text will produce precisely the kinematic numerators for CS theory. Much like SDYM and ZM theory, the CS action has a preferred integer dimension of $D=3$,
\be
S_{\text{CS}} = \int d^3 x \,\text{tr}\left[\epsilon^{\mu\nu\rho} \left(A_\mu \partial_\nu A_\rho + \frac{2}{3}A_{\mu}A_{\nu}A_{\rho}\right)\right]
\label{eq:CSAction}
\ee
where $\epsilon^{\mu\nu\rho}$ is a 3D Levi-Civita symbol. The color-dual properties of this theory were previously studied in Ref.~\cite{Ben-Shahar:2021zww}, which demonstrated that color-kinematics duality is manifest off-shell. \jm{The next 2-3 sentences are new.}
At the classical level, CS is clearly related to $J$-theory (and thus semi-YM) since the CS equation of motion,
\be
\epsilon^{\rho \mu \nu}F_{\mu\nu} =0 \Leftrightarrow F_{\mu\nu}=0 \, ,
\ee
and the Lorenz gauge choice,
\be
\partial_\mu A^\mu=0 \, ,
\ee
are the same as the $J$-theory conditions in \cref{eq:FieldStrengthJ,eq:LorenzGaugeJ}, respectively.

The correspondence between CS and semi-YM can be made more precise at the level of the Feynman rules.
The color-stripped three-point vertex and propagator can be read off from the CS action in \cref{eq:CSAction} producing
\be\label{CSFrule}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} = \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2\varepsilon^{\rho}_3 \qquad \qquad \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho \, .
\ee
At first glance, this seems far removed from the semi-abelian Yang-Mills analogs in \cref{semiYMvert} and \cref{semiYMprop}. After all, semi-YM makes no reference to an anti-symmetric 3-tensor and the mass dimensions of the vertices do not agree. However, to construct the full on-shell amplitudes of CS theory, we must plug in the Fourier transforms of the on-shell currents, $\bar{\varepsilon}(p)$, as described in Ref.~\cite{Ben-Shahar:2021zww}. \jm{How about we use $\varepsilon_\mu$ and $\epsilon_\mu$ like Henrik and Maor?  I'm worried that the bar is getting overused a little.  Or how about $\varepsilon$ and $\tilde{\varepsilon}$ but this will eventually involve $\tilde{\bar{\varepsilon}}$ which is a little gross?} These are related to the polarizations, $\varepsilon_i$, of the Feynman rules above as follows:
\be
\varepsilon^{\mu}(p) = \epsilon^{\mu\nu\rho} p^\nu \bar{\varepsilon}^\rho(p)\, .
\ee
With this definition of the on-shell states, it is clear how the Feynman rules of \cref{CSFrule} are related to those of \cref{semiYMvert} and \cref{semiYMprop} -- we must simply canonicalize the CS kinetic term by absorbing $ \epsilon^{\mu\nu\rho} k^\rho$ into the definition of the three-point vertex:
\be
 \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho
\quad \longrightarrow \quad  \propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}\,.
\ee
This has the effect of orienting the three-point vertex of \cref{CSFrule} by making the replacement $\varepsilon_i \rightarrow \epsilon^{\mu\nu\rho} k_i^\nu \bar{\varepsilon}^\rho_i$ on one of the legs,
\begin{align}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} =  \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2 (\epsilon^{\rho \alpha \beta} k_3^\alpha \bar{\varepsilon}^\beta_3) =    (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)\, ,
\end{align}
where we have used the identity $\epsilon^{\mu\nu\rho}\epsilon^{\rho\alpha\beta} = \eta^{\mu \alpha} \eta^{\nu \beta} -  \eta^{\mu \beta} \eta^{\nu \alpha}$. As desired, the redefined CS vertex is precisely the three-point vertex of semi-YM in \cref{semiYMvert}. Clearly the $D$-dimensional amplitudes of semi-YM can be mapped to the 3D observables of CS theory by making the following replacement on all of the external states of semi-abelian YM:
\be
\varepsilon^\mu_i \rightarrow \epsilon^{\mu\nu\rho} k^\nu_i \bar{\varepsilon}^\rho_i \qquad\qquad  \bar{\varepsilon}_i^\mu \rightarrow \bar{\varepsilon}^\mu_i \, .
\ee
Upon making this replacement, the amplitudes of $D$-dimensional semi-YM are mapped onto 3D CS theory, $\mathcal{A}^{\text{semi-YM}}\rightarrow \mathcal{A}^{\text{CS}}$. While this may seem like a marginal gain at tree-level, by using our $D$-dimensional construction at one-loop, the manifestly color-dual integrands of \cref{SDYMnGon} can resolve regularization ambiguity of CS theory due to its explicit dependence on $D=3$. For more background on CS theory, we refer the reader to Ref.~\cite{Ben-Shahar:2021zww}.

\jm{Do we want to mention something about how it's not obvious in this formulation of CS that the amplitudes vanish?  This also happens for NLSM where it's not obvious why odd multiplicity semi-YM amplitudes with the NLSM external states vanish.}

@@

We use the same conventions as Ref.~\cite{jjmcTASI2014}, which
we quote now. The component-wise definitions of the spinor brackets are
\begin{align}
\langle ab \rangle &= \frac{(a_1 + i a_2)(b_0+b_3)-(b_1 + i b_2)(a_0+a_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\\
[ab] &= \frac{(b_1 - i b_2)(a_0+a_3)-(a_1 - i a_2)(b_0+b_3)}{\sqrt{(a_0+a_3)(b_0+b_3)}}\,,
\end{align}
where the $a_i$ are components of the four-vector,
$k^\mu_a = (a_0,a_1,a_2,a_3)$. For massless momenta, $k_a$ and $k_b$,
one can apply the on-shell condition, $k_a^2=k_b^2=0$, to show that
\begin{equation}
s_{ab} = (k_a+k_b)^2= \langle ab \rangle[ba]\,.
\end{equation}
Four-dimensional polarization dot products with fixed helicity states
are mapped to
\begin{equation}\label{eq:4DPols}
\begin{aligned}
k_a \cdot \varepsilon_b^{(+)} &= \frac{\langle q a \rangle[ab]}{\sqrt{2}\langle q b\rangle}\,,
\qquad\quad \qquad
k_a \cdot \varepsilon_b^{(-)} = -\frac{[qa]\langle ab\rangle}{\sqrt{2}[qb]}\,,
\\
\varepsilon_a^{(-)}\cdot \varepsilon_b^{(+)} &= - \frac{\langle q a\rangle [qb]}{ [qa]\langle q b\rangle} \,,
\qquad \qquad
\varepsilon_a^{(\pm)}\cdot \varepsilon_b^{(\pm)} = 0 \,,
\end{aligned}
\end{equation}
Note that given the above definition, the spinor helicity variables
carry the same mass dimension for both the angle and square
brackets. In terms of the light-cone coordinates that we employed in
\cref{sec:CKLagrangians}, the angles and squares can be rewritten as
\begin{align}
\langle ab \rangle &= \frac{a_w b_u-a_u b_w}{\sqrt{a_ub_u}}\,,
\\
[ab] &= \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{\sqrt{a_ub_u}}\,.
\end{align}
The conventions we use in the text where the holomorphic spinor,
$X(p,k)$, carries extra mass dimension amounts to shifting
the bracket by an overall factor,
\begin{align}
\langle ab \rangle &\rightarrow a_w b_u-a_u b_w \equiv X(a,b)\,,
\\
[ab] &\rightarrow \frac{b_{\bar{w}}a_u-a_{\bar{w}}b_u}{a_ub_u}\equiv Q(a,b)\, .
\end{align}
With this choice we still obtain the same completeness relation for the SDYM spinors,
\begin{equation}
X(a,b)Q(b,a) = s_{ab}\, .
\end{equation}
